Quantum Walks and Application to Quantum Money

### 2025-04-07 main.tex
draft.tex added

@@

\section*{Introduction}

Quantum computing is a revolutionary field that combines quantum mechanics and information theory to redefine computation and information processing. In the latter part of the 20th century, scientists explored the fusion of quantum mechanics and information theory, leading to the birth of quantum information science. This field challenges the classical view of computation by introducing novel concepts like quantum bits (qubits), entanglement, and quantum superposition, which have enabled new algorithms and protocols that outpace their classical counterparts in specific tasks. Classical computers are grounded in bits, which take binary values (0 or 1). Quantum computing introduces the qubit, a fundamental unit of quantum information that can exist in a superposition of states, such as 0 and 1 simultaneously. This foundational difference enables quantum computers to process information in fundamentally new ways. Unlike technologies like DNA computing or optical computing, which describe changes in the physical substrate while retaining classical computational principles, quantum computing changes the computational paradigm itself. A quantum computer uses the principles of quantum mechanics, such as superposition and entanglement, rather than classical mechanics, to process information. In the 1980s, pioneers like Richard Feynman and David Deutsch recognized that certain quantum phenomena could not be efficiently simulated on classical computers. These insights led to the exploration of quantum Turing machines and quantum circuit models, which provided a theoretical framework for quantum computing. The discovery of quantum gates and their role in quantum algorithms formalized the field.

Early quantum algorithms demonstrated that quantum computing could solve certain problems more efficiently than classical methods. Notably, in 1994, Peter Shor introduced a polynomial-time quantum algorithm for integer factorization, which threatened classical cryptographic protocols relying on the difficulty of factoring large integers. Entanglement, a uniquely quantum phenomenon, allows particles to exhibit correlations that defy classical explanation. This property is crucial for many quantum algorithms and protocols, as it enables quantum computers to process and store information in a way that classical computers cannot. Two notable algorithms underscore the potential of quantum computing: Shor’s Algorithm, which efficiently factors integers, undermining RSA encryption and other cryptographic systems based on the hardness of factoring, and Grover’s Algorithm, which provides a quadratic speedup for unstructured search problems, such as searching an unsorted database. These algorithms exemplify the theoretical advantages of quantum computing over classical approaches, even though practical implementations remain challenging.

Quantum systems are fragile and susceptible to decoherence, where quantum states lose their coherence due to interactions with the environment. This makes maintaining quantum states for computation a significant challenge. Quantum error correction methods were developed to address decoherence and other quantum noise. The breakthrough work of Shor and Steane in the mid-1990s introduced error-correction codes that allowed reliable computation despite quantum noise. Building scalable quantum computers requires advances in hardware and experimental techniques. As of now, only small-scale quantum systems with a few qubits have been implemented successfully in laboratories. Quantum computing does not offer universal speedups for all problems. For example, Grover’s algorithm provides only a quadratic speedup for unstructured search, and certain problems remain equally challenging for both quantum and classical computers.

Quantum key distribution protocols, like BB84 and Ekert’s protocol, offer provably secure methods for communication based on the principles of quantum mechanics. Unlike classical cryptography, which relies on computational assumptions, quantum cryptography guarantees security through physical principles. The quantum perspective has provided new insights into classical computing and inspired novel classical algorithms. It has also advanced simulation techniques for quantum systems, benefiting fields like material science and chemistry. Quantum information processing has deepened our understanding of quantum mechanics, shedding light on foundational questions about quantum measurement and entanglement. For example, experiments testing Bell’s inequalities have confirmed the non-classical correlations predicted by quantum theory.

While practical quantum computing is still in its infancy, significant progress has been made. Quantum hardware companies like IBM, Google, and Rigetti have developed small-scale quantum processors capable of performing limited computations. Platforms such as Qiskit and Cirq enable researchers to experiment with quantum programming and algorithm development. Efforts to build scalable, fault-tolerant quantum computers are ongoing, alongside investigations into alternative models of quantum computation, such as topological and cluster-state quantum computing. Despite these advances, many open questions remain about the scope and ultimate power of quantum computation. While quantum computers will not replace classical ones for all tasks, they promise to revolutionize fields where their unique capabilities provide exponential speedups or new forms of computation.

Quantum computing represents a profound shift in how we understand and leverage computation. By replacing classical mechanics with quantum mechanics as the foundation for processing information, quantum computing has opened new avenues for scientific discovery and technological innovation. While challenges persist, the theoretical and experimental advances made thus far underscore the transformative potential of this exciting field.

\section*{Foundations of Quantum Computation}

\section*{Understanding a Qubit}
A qubit (quantum bit) is the fundamental unit of quantum information, analogous to a classical bit. However, unlike a classical bit, which can only exist in one of two definite states ($0$ or $1$), a qubit can exist in a linear combination, or superposition, of both states simultaneously. Mathematically, the state of a single qubit is represented as:

\begin{equation}
|\psi\rangle = \alpha |0\rangle + \beta |1\rangle
\end{equation}

Here:
\begin{itemize}
    \item $|0\rangle$ and $|1\rangle$ are the basis states (analogous to $0$ and $1$ in classical computing).
    \item $\alpha$ and $\beta$ are complex numbers known as probability amplitudes.
\end{itemize}

The normalization condition \(|\alpha|^2 + |\beta|^2 = 1\) ensures that the probabilities of measuring the qubit in either the $|0\rangle$ or $|1\rangle$ state sum to 1.

The qubit's unique ability to exist in a superposition of states is what gives quantum computers their immense computational potential, enabling them to process and store information in fundamentally different ways than classical computers.

\section*{Superposition}
Superposition is a quantum phenomenon where a qubit exists in a combination of multiple states simultaneously. For example, a qubit in the state:

\begin{equation}
|\psi\rangle = \frac{1}{\sqrt{2}}|0\rangle + \frac{1}{\sqrt{2}}|1\rangle
\end{equation}

is in an equal superposition of $|0\rangle$ and $|1\rangle$. This means that if we measure the qubit, there is an equal probability (50\%) of observing it in either state.

Key properties of superposition:
\begin{itemize}
    \item \textbf{Parallelism:} When a quantum system is in superposition, it can perform computations for all possible states simultaneously. For instance, a single qubit can encode two states ($|0\rangle$ and $|1\rangle$) at the same time, while $n$ qubits can encode $2^n$ states.
    \item \textbf{Measurement Collapse:} When a qubit in superposition is measured, it collapses into one of its basis states, $|0\rangle$ or $|1\rangle$, with a probability given by $|\alpha|^2$ and $|\beta|^2$, respectively.
\end{itemize}

Superposition allows quantum systems to explore multiple possibilities in parallel, which is critical for quantum algorithms such as Grover's search or Shor's factoring.

\section*{Entanglement}
Entanglement is a uniquely quantum phenomenon where two or more qubits become correlated in such a way that the state of one qubit is directly related to the state of the other, regardless of the physical distance between them. When qubits are entangled, the measurement of one qubit instantly determines the state of the other.

An example of an entangled state for two qubits is the Bell state:

\begin{equation}
|\psi\rangle = \frac{1}{\sqrt{2}} (|00\rangle + |11\rangle)
\end{equation}

Here:
\begin{itemize}
    \item $|00\rangle$ means both qubits are in the $|0\rangle$ state.
    \item $|11\rangle$ means both qubits are in the $|1\rangle$ state.
\end{itemize}

If one qubit is measured to be $|0\rangle$, the other qubit will instantly collapse to $|0\rangle$, and similarly for $|1\rangle$, regardless of the distance between them.

Key features of entanglement:
\begin{itemize}
    \item \textbf{Non-Local Correlations:} Entanglement defies classical intuition, as it suggests that measurement outcomes are correlated even across vast distances, a phenomenon supported by experiments validating Bell's Theorem.
    \item \textbf{Applications:} Entanglement is a resource for many quantum technologies, including quantum teleportation, quantum cryptography (e.g., secure communication via the BB84 protocol), and quantum error correction.
\end{itemize}

Entanglement, along with superposition, forms the backbone of quantum computing and quantum communication, enabling capabilities that are impossible in the classical world.

\subsection*{Quantum Gates and Circuits}

Quantum logic gates are the fundamental building blocks of quantum circuits, manipulating qubits—the basic units of quantum information. Unlike classical logic gates, quantum gates are reversible and represented by unitary matrices, ensuring the preservation of quantum information.

\subsection*{Key Quantum Logic Gates}

\subsubsection*{1. Pauli-X Gate (NOT Gate)}
\textbf{Operation:} Flips the state of a qubit from $|0\rangle$ to $|1\rangle$ and vice versa. \\
\textbf{Matrix Representation:}
\[
X = 
\begin{bmatrix}
0 & 1 \\
1 & 0
\end{bmatrix}
\]

\subsubsection*{2. Hadamard Gate (H Gate)}
\textbf{Operation:} Creates a superposition state, transforming $|0\rangle$ into $\frac{|0\rangle + |1\rangle}{\sqrt{2}}$ and $|1\rangle$ into $\frac{|0\rangle - |1\rangle}{\sqrt{2}}$. \\
\textbf{Matrix Representation:}
\[
H = \frac{1}{\sqrt{2}}
\begin{bmatrix}
1 & 1 \\
1 & -1
\end{bmatrix}
\]

\subsubsection*{3. Controlled-NOT Gate (CNOT Gate)}
\textbf{Operation:} A two-qubit gate that flips the state of the target qubit if the control qubit is in the state $|1\rangle$. \\
\textbf{Matrix Representation:}
\[
CNOT =
\begin{bmatrix}
1 & 0 & 0 & 0 \\
0 & 1 & 0 & 0 \\
0 & 0 & 0 & 1 \\
0 & 0 & 1 & 0
\end{bmatrix}
\]

\subsubsection*{4. Toffoli Gate (CCNOT Gate)}
\textbf{Operation:} A three-qubit gate that flips the state of the target qubit if both control qubits are in the state $|1\rangle$. \\
\textbf{Matrix Representation:}
\[
CCNOT =
\begin{bmatrix}
1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 \\
0 & 0 & 0 & 0 & 0 & 0 & 1 & 0
\end{bmatrix}
\]

\subsection*{Universal Gates}
A set of gates is considered \textbf{universal} if any unitary operation can be approximated to arbitrary precision using a sequence of gates from this set. For example, single-qubit rotation gates combined with the CNOT gate form a universal set, enabling the construction of any quantum algorithm.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction to Quantum Walks}

The concept of a \textit{quantum walk} is a quantum analog of the classical random walk and plays a significant role in quantum algorithms. There are two main types of quantum walks: \textit{continuous-time} and \textit{discrete-time} quantum walks, both of which exhibit behaviors that are significantly different from classical random walks.

\section*{Continuous-Time Quantum Walks}

Classical continuous-time random walks (Markov chains) are a foundational concept in stochastic processes, where time is treated as a continuous variable. In this framework, a walker moves from one vertex of a graph to an adjacent vertex at any time, with probabilities evolving over time. The probability of the walker being at a vertex can be visualized as a liquid seeping from one vertex to its neighbors. Initially, the walker is most likely to be found at the starting vertex, but as time progresses, the probability shifts to neighboring vertices. This dynamic is governed by a constant transition rate \( \gamma \), uniform across all vertices and time.

To model this mathematically, the probability of transitioning from one vertex to another in an infinitesimal time interval \( \Delta t \) is proportional to \( \gamma \Delta t \). For a vertex \( x_j \) with degree \( d_j \) (the number of neighboring vertices), the probability of the walker staying on \( x_j \) is

\[
1 - d_j \gamma \Delta t,
\]

while the probability of transitioning to a neighboring vertex is

\[
\gamma \Delta t.
\]

The evolution of probabilities over time is captured by the transition matrix \( M(t) \), where the entry \( M_{ij}(t) \) represents the probability of transitioning from vertex \( x_j \) to vertex \( x_i \) in time \( t \). The dynamics of the system are described by a differential equation derived from the transition matrix and an auxiliary matrix \( H \), known as the generating matrix. This equation,

\[
\frac{dM(t)}{dt} = -H M(t),
\]

has a solution expressed as

\[
M(t) = e^{-Ht},
\]

which allows for the computation of probability distributions at any given time. The solution is verified using the Taylor series expansion of the exponential function.

The passage from classical continuous-time Markov chains to quantum walks involves a standard quantization process. In classical random walks, probabilities evolve via the transition matrix \( M \). In contrast, the quantization process replaces the vector of probabilities with a state vector (probability amplitudes) and the non-unitary transition matrix with a unitary evolution operator, as required by quantum mechanics. This involves transforming the generating matrix \( H \) into a Hermitian operator and then constructing a unitary operator

\[
U(t) = e^{-iHt}
\]

by multiplying \( H \) with the imaginary unit \( i \).

In a continuous-time quantum walk, the quantum state evolves over time according to the unitary operator \( U(t) \). If the initial state of the system is \( |\psi(0)\rangle \), the state at time \( t \) is given by

\[
|\psi(t)\rangle = U(t) |\psi(0)\rangle.
\]

The probability distribution for observing the walker at a specific vertex \( k \) is computed as

\[
p_k = |\langle k | \psi(t) \rangle|^2,
\]

where \( |k\rangle \) represents the state corresponding to vertex \( k \).

The continuous-time quantum walk introduces several key differences from its classical counterpart. Unlike classical random walks, which rely on probabilities, quantum walks involve probability amplitudes, enabling interference effects. These effects allow quantum walks to explore the graph more efficiently, leading to applications in quantum algorithms with speedups over classical approaches. Additionally, the transition matrix in classical random walks is replaced with a Hermitian operator in the quantum case, ensuring unitary evolution.

*********Examples and further discussions here*********

% Overall, the transition from classical continuous-time Markov chains to continuous-time quantum walks highlights fundamental differences between classical and quantum dynamics, particularly the role of interference and the constraints of unitary evolution in quantum systems. These differences underscore the potential of quantum walks in quantum computing and algorithm design.

\section*{Discrete-Time Quantum Walks}
*******************
The discrete-time coined quantum walk (DTQW) is a quantum analog of classical random walks, leveraging the principles of quantum mechanics to achieve unique computational and dynamical properties. It is defined on a graph or lattice, with the walker's state described by both its position and an internal "coin" state. The walker's position \( n \) on a one-dimensional lattice (line) is represented by a vector \( |n\rangle \) in an infinite-dimensional Hilbert space \( \mathcal{H}_P \), where the computational basis for \( \mathcal{H}_P \) is

\[
\{ |n\rangle : n \in \mathbb{Z} \}.
\]

The walker's internal state, or "coin," is represented by a two-dimensional Hilbert space \( \mathcal{H}_C \) with basis states \( |0\rangle \) and \( |1\rangle \). The coin state determines the direction of the walker's movement, and the total state of the system resides in the tensor product space

\[
\mathcal{H} = \mathcal{H}_C \otimes \mathcal{H}_P,
\]

combining the coin and position spaces.

The dynamics of the walk are governed by two key operators: the coin operator and the shift operator. The coin is represented by a unitary matrix \( C \) acting on \( \mathcal{H}_C \). Common choices include the Hadamard matrix, which creates superpositions of the coin states. The Hadamard coin operator \( H \) is defined as

\[
H = \frac{1}{\sqrt{2}} \begin{pmatrix} 1 & 1 \\ 1 & -1 \end{pmatrix}.
\]

The shift operator \( S \) moves the walker based on the coin state. For example, if the coin state is \( |0\rangle \), the walker moves to \( |n+1\rangle \), and if the coin state is \( |1\rangle \), the walker moves to \( |n-1\rangle \). The shift operator is explicitly given by

\[
S = \sum_{n} |n+1\rangle \langle n| \otimes |0\rangle \langle 0| + |n-1\rangle \langle n| \otimes |1\rangle \langle 1|.
\]

The evolution of the system is described by the unitary operator

\[
U = S \cdot (I \otimes C),
\]

where \( I \) is the identity operator on \( \mathcal{H}_P \). The state of the walker at time \( t+1 \) is given by

\[
|\psi(t+1)\rangle = U |\psi(t)\rangle.
\]

Consider a quantum walk on a one-dimensional lattice. The walker's state evolves in superposition, allowing it to explore multiple paths simultaneously. Interference between these paths leads to faster spreading compared to classical random walks.

Key features of DTQW include superposition, where the walker exists in a superposition of positions and coin states, enabling parallel exploration of multiple paths; interference, where quantum interference enhances or suppresses certain paths, leading to non-classical behavior; and quadratic speedup, where quantum walks spread faster than classical random walks, often quadratically, which is advantageous for algorithmic applications.

DTQWs are used in quantum algorithms, such as quantum search and graph traversal, and in simulating physical systems like energy transport in photosynthesis. They also serve as a framework for universal quantum computation. The discrete-time quantum walk is a powerful tool in quantum computing, offering a rich interplay between quantum mechanics and graph theory. Its ability to exploit superposition and interference enables applications in algorithm design, simulation, and computation, making it a cornerstone of quantum information science.

*********Examples and further discussions here*********

\subsection*{Comparison of Models}

>>>>

\newpage
\section*{Abstract}

This thesis explores the foundations of quantum computation, focusing on quantum walks and their application to quantum money. Quantum walks, particularly continuous-time quantum walks based on group actions, serve as a powerful computational tool with applications in search algorithms and cryptographic protocols. We examine their mathematical structure and their advantages over classical random walks, emphasizing their efficiency in state evolution and probability distribution spreading. As a part of this work, we examine efficient implementations of transforms such as the Quantum Fourier Transform (QFT) and the Quantum Hartley Transform (QHT), analyzing their role in encoding quantum states for secure cryptographic applications. In particular, we discuss a novel instantiation of a quantum money scheme based on QHT, leveraging its unique properties for improved security and efficiency.
To ensure the robustness of this quantum money scheme, we develop a verification mechanism utilizing quantum walks. Unlike previous approaches, which rely on standard quantum state measurements, our method employs continuous-time quantum walks to authenticate quantum money, preventing counterfeiting while maintaining computational feasibility. Additionally, we present a detailed discussion on the efficient implementation of this scheme, including optimized circuit designs and error mitigation strategies.

@@

\chapter{Introduction}
Quantum money, first introduced in a seminal paper by Wiesner [29], is a form of currency represented by quantum states. Unlike classical money, an ideal quantum bill cannot be counterfeited due to the no-cloning theorem of quantum mechanics, which prohibits making an identical copy of an unknown quantum state. Wiesner’s original scheme (now termed a \emph{private-key quantum money} scheme) had significant practical drawbacks: it required the issuing bank’s secret key to verify each banknote, meaning the bank had to be involved in every transaction. This reliance on the bank for verification severely limits the usability of private-key quantum money schemes.

In 2009, Aaronson [1] proposed the first \emph{public-key quantum money} scheme, in which anyone can verify a quantum banknote while only the bank can mint (issue) new valid banknotes. Public-key quantum money removes the transactional bottleneck of involving the bank every time. Aaronson’s specific scheme was later broken by Lutomirski et al.~[24]. In the years since, several alternative public-key quantum money constructions have been explored [2, 17, 30, 19, 20, 23, 31]. However, each of these proposals has either been broken by further cryptanalysis [14, 26, 7, 23] or relies on unconventional cryptographic assumptions, making their security or practicality questionable.

\medskip\noindent\textbf{Quantum Money from Group Actions and the Fourier Transform.} 
A promising candidate for secure public-key quantum money based on more standard assumptions was recently proposed by Zhandry [31]. In Zhandry’s scheme, the core idea is to use an abelian group action as the foundation for the quantum state. Each money state (quantum banknote) is prepared as a \emph{group-action Fourier state}, and the corresponding serial number is an element of the underlying group (we will elaborate on these terms in later chapters). The verification algorithm in this scheme uses a group-action phase kickback routine to extract the serial number from the quantum banknote and check its validity. Doliskani [15] later proved that Zhandry’s scheme is secure in the generic group action model, providing evidence for its security under well-defined assumptions.

\medskip\noindent\textbf{This Work – Motivation and Contributions.} 
In this thesis, we adapt Zhandry’s group-action quantum money scheme by replacing its use of the Quantum Fourier Transform with the \emph{Quantum Hartley Transform} (QHT) over finite abelian groups. The motivation behind this substitution is multifold. First, using the Hartley transform causes the banknotes to have \emph{real} amplitudes instead of complex amplitudes. We believe this shift from complex to real quantum states may offer both computational and theoretical advantages. For example, certain mathematical identities hold for any two real orthonormal bases that do not generally hold for complex bases; such differences can affect the properties of quantum states and were a barrier in prior analyses (indeed, an attempted “quantum lightning” construction in [31] failed due to properties of complex phases that do not arise with real amplitudes). Beyond this specific context, our work is a first step toward employing real-valued quantum transforms in quantum cryptography and quantum algorithms. To the best of our knowledge, this is the first significant quantum cryptographic construction that makes essential use of the quantum Hartley transform. We hope that exploring real-amplitude quantum states will inspire further research into new quantum algorithms and optimizations for real-valued transforms.

We now summarize our main contributions:
\begin{itemize}
    \item We propose a new \textbf{public-key quantum money scheme} based on abelian group actions and the quantum Hartley transform. This scheme is an adaptation of Zhandry’s group-action money scheme, modified to output real-amplitude quantum states. 
    \item We identify a breakdown in the original verification procedure when the Fourier transform is naively replaced by the Hartley transform: the Hartley-based verification fails to distinguish certain illegitimate banknotes. To overcome this, we design a \textbf{new verification algorithm} that relies on applying additional group action “twists” and utilizes quantum walks. This new verification algorithm successfully differentiates valid banknotes from forgeries despite the use of real amplitudes.
    \item We show that the serial number associated with a money state (which is a hidden group element) can be efficiently computed from the quantum state. In particular, we develop a \textbf{continuous-time quantum walk algorithm} to extract the serial number of a given banknote state. This algorithm leverages the structure of the group action and the spectral properties of an associated Cayley graph. Our method demonstrates that one can recover the hidden group element (serial) with high success probability, which not only underpins the new verification technique but is also of independent interest.
    \item In the course of our construction, we introduce a new algorithm for efficiently implementing the \textbf{quantum Hartley transform} and related real transforms. We present a recursive technique that exploits the structure of the Hartley transform, yielding a lower gate complexity than prior approaches in the literature. We also illustrate how other real-valued quantum transforms (e.g., the quantum sine transform) can be implemented using the Hartley transform as a subroutine. These results contribute to the broader goal of optimizing real quantum transforms for practical use.
\end{itemize}

In summary, our work combines ideas from quantum cryptography (public-key quantum money), quantum algorithms (quantum walks), and quantum signal processing (Hartley and related transforms) to create a novel and potentially practical quantum money scheme.

\medskip\noindent\textbf{Thesis Organization.} 
The remainder of this thesis is organized as follows. In **Chapter 2**, we provide background on the basics of quantum computation and the theory of group actions, including the concept of cryptographic group actions which underlie our money scheme. **Chapter 3** covers quantum walks, both continuous-time and discrete-time, outlining their differences and importance in quantum algorithms. In **Chapter 4**, we apply the quantum walk framework to group actions: we describe how a continuous-time quantum walk on a group action can be simulated efficiently, a crucial step for our serial number extraction algorithm. **Chapter 5** details the quantum money scheme itself and the role of the Hartley transform in its construction. We explain how the scheme is formulated and discuss the advantages and challenges introduced by using the Hartley transform. Finally, **Chapter 6** presents the verification algorithm for the Hartley-based quantum money. This chapter shows how we use quantum walks (as developed in Chapter 4) to reliably verify banknotes and compute their serial numbers, thereby addressing the verification issues that arose from the use of real amplitudes.
\vspace{1em}

\chapter{Background on Quantum Computation and Group Actions}

\section{Quantum Computation Basics}
Quantum computation operates on information stored in quantum states. The fundamental unit of quantum information is the \emph{qubit}, which, unlike a classical bit, can exist in a superposition of basis states. A qubit is described by a state $|\psi\rangle = \alpha |0\rangle + \beta |1\rangle$ (in Dirac notation), where $\alpha,\beta \in \mathbb{C}$ and $|\alpha|^2 + |\beta|^2 = 1$. A system of $n$ qubits resides in a $2^n$-dimensional Hilbert space and its state can be a superposition of all $2^n$ basis basis states $|b_1b_2\cdots b_n\rangle$.

Computation is carried out by unitary transformations on these states. A quantum algorithm is typically a sequence of quantum gates (unitary operators) applied to an initial state (usually a simple basis state like $|00\cdots0\rangle$), followed by a measurement of some qubits to obtain a classical result. A crucial aspect of quantum computation is that certain transformations—such as the Fourier transform over an $N$-element group—can be implemented very efficiently on a quantum computer (in $\operatorname{poly}(\log N)$ time), whereas their classical counterparts might require $\operatorname{poly}(N)$ time.

One example important to this work is the \textbf{quantum Fourier transform} (QFT). For an abelian group $G$ of order $N$, the QFT is the unitary map $|g\rangle \mapsto \frac{1}{\sqrt{N}}\sum_{h \in G} \chi_h(g)\, |h\rangle$, where $\{\chi_h\}$ are the characters (complex exponential homomorphisms) of $G$. In more familiar terms, when $G = \mathbb{Z}_N$ (the integers mod $N$), the QFT takes $|j\rangle$ to $\frac{1}{\sqrt{N}}\sum_{k=0}^{N-1} e^{2\pi i j k / N} |k\rangle$. The QFT is a key component in many quantum algorithms, including Shor’s factoring algorithm and certain quantum money schemes (e.g., Zhandry’s) that rely on hiding information in phase relationships. The QFT can be implemented using $O((\log N)^2)$ basic gates for many groups [13], making it an efficient operation.

Another concept we will use is the \textbf{quantum Hartley transform} (QHT). The Hartley transform is an alternative to the Fourier transform that produces only real values (in the classical case) by using sinusoidal basis functions (sums of sine and cosine) instead of complex exponentials. The \emph{quantum} Hartley transform is the unitary transform corresponding to the classical Hartley transform. For example, the Hartley transform on $\mathbb{Z}_N$ maps a basis state $|j\rangle$ to $\frac{1}{\sqrt{N}}\sum_{k=0}^{N-1} \mathrm{cas}\!\Big(\frac{2\pi j k}{N}\Big)\, |k\rangle$, where $\mathrm{cas}(x) = \cos x + \sin x$ is the Hartley kernel. Because $\cos x + \sin x$ is real for all $x$, the resulting amplitudes are real numbers. We will discuss the Hartley transform in more detail in Chapter 5, including how to implement it efficiently on a quantum computer.

A final fundamental principle to note is the \textbf{no-cloning theorem}, which underpins the security of quantum money. This theorem states that there is no physical operation that can make a copy of an arbitrary unknown quantum state. This contrasts sharply with classical information, which can be copied freely. The no-cloning theorem means that an adversary cannot duplicate a quantum banknote without fundamentally disturbing it – any attempt to “clone” the state will fail unless the adversary already has complete classical information describing that state.

Throughout this thesis, when we refer to an algorithm as \emph{efficient} or \emph{feasible}, we typically mean it can be performed by a quantum polynomial-time (QPT) algorithm. This is analogous to classical polynomial time, but on a quantum computer. All of the algorithms we propose (minting, verification, quantum walk simulation, etc.) are intended to be efficient in this sense.

\section{Group Actions in Cryptography}
At the heart of our quantum money scheme is the notion of a group action. A \textbf{group action} of a group $G$ on a set $X$ is a function $*\colon G \times X \to X$ that satisfies two properties: (1) the identity element $e \in G$ acts trivially on every $x \in X$ ($e * x = x$), and (2) the action is compatible with the group operation, meaning $g*(h* x) = (gh)*x$ for all $g,h \in G$ and $x \in X$. We will often denote the result of $g$ acting on $x$ by $g*x$ or simply $g x$ when the action is clear from context.

The orbit of an element $x \in X$ under the action is the set $G * x = \{g*x : g \in G\} \subseteq X$. If there is exactly one orbit (i.e., $G * x = X$ for some $x$), the action is called \emph{transitive}. If, in addition, no non-identity group element fixes any $x$ (i.e., $g * x = x$ implies $g = e$), then the action is \emph{free}. A group action that is both transitive and free is sometimes called a \textbf{regular action}. In a regular action, for any two elements $y, z \in X$, there is a unique group element $g$ such that $g * y = z$. In other words, $G$ acts like a permutation group that moves elements of $X$ around, and knowing the starting and ending point uniquely determines the “move” $g$. In this case $|X| = |G|$, and we can think of $X$ as essentially a copy of $G$ with the group structure “hidden” by the action.

Group actions are ubiquitous in modern cryptography. A \textbf{cryptographic group action} is one where the action is easy to compute in the forward direction, but inverting the action (recovering $g$ from $x$ and $g*x$) is computationally hard. This essentially functions as a one-way function: given $x$ and $g$, it is easy to compute $y = g*x$, but given $x$ and $y$ it is infeasible to find $g$ (except with negligible probability). One classical example is the discrete logarithm problem: consider the cyclic group $G = \mathbb{Z}_p^*$ (the multiplicative group of a prime field) acting on itself by exponentiation. Let $X = G$ and define $g * x = x^g \pmod p$ (treating group elements as integers). Here $g$ is an exponent. This is a group action (in fact, just the group’s own operation written differently), and computing $x^g \bmod p$ is efficient. However, given $x$ and $y = x^g$, finding $g$ requires solving a discrete log problem, which is believed to be hard for suitable choices of $p$. Thus, exponentiation can be viewed as a one-way group action.

Another prominent example comes from elliptic-curve cryptography: the action of an ideal class group on a set of elliptic curves. In schemes like CSIDH (Commutative Supersingular Isogeny Diffie–Hellman), an abelian group of ideal classes $G$ acts on the set $X$ of supersingular elliptic curves by isogeny (an isogeny is a structure-preserving map between curves). Starting from a curve $x_0$, one can efficiently compute $g * x_0$ (which is another elliptic curve) for any group element $g$, but given two curves $x_0$ and $x = g * x_0$, it is believed to be computationally hard to recover $g$. This hard problem is the foundation of isogeny-based cryptography. We refer to such $G$ and $X$ as a \emph{cryptographic group action} pair.

Cryptographic group actions provide a natural platform for public-key quantum money schemes. In such schemes, the secret “serial number” of a quantum banknote can be an element $g \in G$, and the quantum state of the banknote can be some function of $g*x_0$ for a public base point $x_0 \in X$. The hardness of inverting the action ($x_0$ and $x = g*x_0$ do not easily reveal $g$) contributes to the security against counterfeiting, while the structure of the group action can be used to efficiently verify authenticity (as we will see with the Fourier or Hartley transform techniques).

We will assume throughout that we have a family of abelian group actions $(G,X)$ that are cryptographically suitable: $|G|$ is large (growing with the security parameter), the action is efficiently computable, transitive and free (so $|X|=|G|$), and inverting the action is infeasible without the secret key. Such families are conjectured to exist (for example, the isogeny-based actions, and others discussed in cryptographic literature).

Finally, we note that working in the \textbf{generic group model} (or generic group \emph{action} model) is a common theoretical approach to analyze security. In a generic model, the group elements are treated as black-box labels without exploiting any special algebraic structure beyond group axioms. Doliskani’s proof of security for Zhandry’s scheme [15] was in such a generic model for group actions, lending confidence that no “generic” attack can break the scheme. In this thesis, we focus on designing the scheme and algorithms at a high level and assume the underlying group action is secure; a detailed security proof is beyond our scope, though we rely on prior results for reassurance.

\chapter{Quantum Walks – Continuous and Discrete}
Random walks on graphs are a powerful tool in classical algorithms, and their quantum analogues, called \emph{quantum walks}, play a fundamental role in many quantum algorithms. Like classical random walks, quantum walks come in two flavors: \textbf{continuous-time} and \textbf{discrete-time}. Both types of quantum walks can exhibit dramatically different behavior from their classical counterparts due to quantum interference, which can be leveraged for algorithmic speedups in various problems (for example, quantum walks have led to faster algorithms for database search and element distinctness [4, 5], as well as exponential speedups in oracular problems [10]).

\section{Continuous-Time Quantum Walks}
A continuous-time quantum walk (CTQW) takes place on the vertices of a graph and is defined by a Hamiltonian (energy operator) that drives the evolution of a quantum state across the graph. Formally, consider a graph $\Gamma$ with vertex set $V$ of size $n$. We associate a basis state $|v\rangle$ with each vertex $v \in V$. In a continuous-time \emph{classical} random walk, one can describe the probability distribution $p(t)$ over vertices at time $t$ by a differential equation $dp/dt = -L p(t)$, where $L$ is the Laplacian matrix of the graph. By analogy, in a continuous-time \emph{quantum} walk, we use the Schrödinger equation:
\[ i\,\frac{d}{dt}|\psi(t)\rangle = H\,|\psi(t)\rangle, \] 
where $H$ is the Hamiltonian of the walk and $|\psi(t)\rangle$ is the state of the quantum walk at time $t$. The formal solution is $|\psi(t)\rangle = e^{-i H t} |\psi(0)\rangle$. Intuitively, $H$ encodes the connectivity of the graph, and the quantum state evolves continuously, “exploring” the graph’s vertices.

Typically, for a quantum walk on an undirected graph, a natural choice of Hamiltonian $H$ is the adjacency matrix of the graph (sometimes the Laplacian or a normalized version is used, but the adjacency matrix suffices for our purposes). If $A$ is the adjacency matrix of $\Gamma$, then $H = A$ means that the amplitudes flow between adjacent vertices over time. Specifically, if $|v\rangle$ and $|w\rangle$ are connected by an edge, the Hamiltonian will induce oscillations of amplitude between $|v\rangle$ and $|w\rangle$. Because $H$ is Hermitian, $e^{-iA t}$ is unitary, ensuring the evolution is coherent.

One notable feature of CTQWs is that interference effects can cause the walker to propagate through the graph in ways that a classical random walk cannot. For example, quantum walkers can have ballistic spreading (moving roughly linearly fast with time) on certain graphs, whereas classical diffusion is much slower (like $\sqrt{t}$ spread). These effects underlie algorithmic speedups. For instance, the continuous-time walk on a certain $N$-vertex graph can reach a target vertex in $O(N^{1/2})$ time, whereas a classical random walk might take $O(N)$ steps to achieve the same, illustrating a quadratic speedup in some search problems.

\section{Discrete-Time Quantum Walks}
Discrete-time quantum walks proceed in a sequence of steps rather than evolving continuously. In the discrete-time case, we typically require an additional “coin” space or a similar construction to allow the walker to move without violating unitarity (since a deterministic move on an undirected graph can lead to stagnation without a coin toss to break symmetry). In the \textbf{coined model} of a quantum walk, the state is a tensor product $|\text{position}\rangle \otimes |\text{coin}\rangle$. At each step, a unitary coin flip (like a Hadamard gate generalizing a two-sided coin to a multi-sided one for vertices of higher degree) is applied to the coin register, and then a conditional shift moves the walker to a new vertex based on the coin outcome. This two-part operation (coin flip + shift) constitutes one step of the walk, which is a unitary transformation on the overall Hilbert space.
